prolfqua - An R-package for Analysing Proteomics Label Free Quantification Experiments

### 2022-04-05 inst/MyArticle/paper.Rmd
added dataset description.

TODO(wew): Ionstar dataset, give Pride ids, describe MaxQuant settings and version. Give reference to Experiment hub and the _prolfquaData_ package.

For this benchmark dataset, the fraction of regulated proteins (_E. coli_) is relatively large. Therefore, to calibrate the measurement, we used human proteins only. In practice, it is unknown which proteins do and which don't change. Therefore, the assumption used when normalizing, scaling the data, and thus calibrating the measurement is that only a small subset of the proteins in the dataset is differentially regulated.

@@

The peptide intensities from the _MaxQuant_ _peptide.txt_ file were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only. We removed _one hit wonders_, i.e. proteins with a single peptide assignment. No further data filtering was used. We also did not use imputation. The protein level linear model was fitted to the protein intensities inferred from the peptide intensities using the Tukey's median polish. The mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model were fitted to the peptide intensities. The details of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` benchmark implementation are available online  [BenchmarkingIonstarData](https://wolski.github.io/prolfqua/articles/BenchmarkingIonstarData.html).

For benchmarking of `r BiocStyle::Biocpkg("MSstats")` we used the defaults specified in the [Benchmarking_MSstats](https://wolski.github.io/prolfqua/articles/)

@@

>>>>

To evaluate the performance of differential expression analysis, we use the _IonStar_ benchmark dataset[@shen2018ionstar], available from the Proteomics Identifications Database (PRIDE) identifier PXD003881. $DH5\alpha$ _E. coli_ lysate was spiked in human pancreatic cancer cells(Panc-1) lysate at five different levels: $3\%$ _E. coli_, $4.5\%$ _E. coli_, $6\%$ _E. coli_, $7.5\%$ _E. coli_ and $9\%$ _E. coli_. We annotated these dilutions from smallest to largest with the letters a, b, c, d, e. We obtain different fold changes by comparing the various dilutions, e.g., when comparing dilution _e_ (9%) against dilution _d_ (7.5%), the expected fold change is 1.2 for E. coli proteins and 1 for human proteins.
We processed the raw data using MaxQuant [@Cox2008MaxQuant] version Version 1.6.10.43, with MaxQuant default settings. The text files generated by Maxquant are available in the _prolfquadata_ R package [@prolfquadata].

@@

@@

The peptide intensities from the _MaxQuant_ _peptide.txt_ file were $log_2$ transformed, and subsequent robust z-score transformation, where median and the mean average deviation (mad) were obtained from the human proteins only. We removed _one hit wonders_, i.e. proteins with a single peptide assignment. No further data filtering was used. We also did not use imputation. The protein level linear model was fitted to the protein intensities inferred from the peptide intensities using the Tukey's median polish. The mixed effect linear model, and the `r BiocStyle::Biocpkg("PECA")` model were fitted to the peptide intensities. The details of the `r BiocStyle::Githubpkg("fgcz/prolfqua")` benchmark implementation are available online  [BenchmarkingIonstarData](https://wolski.github.io/prolfquabenchmark/articles/BenchmarkingIonstarData.html).

For benchmarking of `r BiocStyle::Biocpkg("MSstats")` we used the MSstats default parameters (for details see: [Benchmarking_MSstats](https://wolski.github.io/prolfquabenchmark/articles/))
